---
title: "Hypothesis Testing in Linear Models"
format: html
code-link: true
editor_options: 
  chunk_output_type: inline
toc: true
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE
)
```

## Incremental contribution of an explanatory variable

We want to test whether `experience` has any significant impact on the wage apart from `education`.

```{r}
library(AER)
data("CPS1985")
cps85 <- CPS1985

wage_model1 <- lm(log(wage) ~ education, data = cps85)
wage_model2 <- lm(log(wage) ~ education + experience, data = cps85)

anova(wage_model1, wage_model2)

```

Check

```{r}
anova(wage_model1) |> names()
rss_r <- anova(wage_model1)$`Sum Sq`[2]
rss_ur <- anova(wage_model2)$`Sum Sq`[3]
m <- 1; n <- nrow(cps85); k <- 3

F_cal <- ((rss_r - rss_ur)/m)/((rss_ur)/(n - k)); F_cal
F_cri <- qf(p = 0.05, df1 = m, df2 = n - k)# Calculate critical ot table F value
F_cri

```

## Testing equality between two regression coefficients

whether the impact of education and experience are the same or not i.e. $H_0: \beta_1 = \beta_2$ or $\beta_1 - \beta_2 = 0$

```{r}
linearHypothesis(wage_model2, "education - experience = 0")
```

Now, if $\beta_1 > \beta_2$ we can infer that the contribution of `education` is more than that of `experience`, and vice-versa.

```{r}
summary(wage_model2)
```

## Testing the validity of linear restriction

We use the Restricted Least Square Approach.

```{r}
library(Ecdat)
data(Metal)
str(Metal)
metal <- Metal
```

Our log-log model is: $$
log(Y_i) = \beta_0 + \beta_1log(L_i) + \beta_2log(K_i)+ u_i
$$

where, $H_0:\beta_1 + \beta_2 = 1$ $\implies$ Constant Returns to Scale

So, we want to check whether the production function exhibits CRS?

```{r}
# unrestricted  model
metal_model1 <- lm(log(va) ~ log(labor) + log(capital), data = metal)

# restricted model
metal_model2 <- lm(log(va/capital) ~ log(labor/capital) , data = metal)

# F stat calculation and comparison
anova(metal_model1) |> names()
rss_r <- anova(metal_model2)$`Sum Sq`[2]
rss_ur <- anova(metal_model1)$`Sum Sq`[3]
m <- 1; n <- nrow(metal); k <- 3

F_cal <- ((rss_r - rss_ur)/m)/((rss_ur)/(n - k)); F_cal
F_cri <- qf(p = 0.1, df1 = m, df2 = n - k)# Calculate critical ot table F value
F_cri
```

## Structural Break Analysis using the F-stat (also known as the Chow test)

Let's look at the data first.

```{r}
# get the time series data
data("USMacroG", package = "AER")
class(USMacroG)

# convert the data type to tsibble
library(tsibble)
quarterly_us_macro <- as_tsibble(USMacroG)

# convert the quarterly data to yearly data
yearly_us_macro <- quarterly_us_macro |>
  pivot_wider(names_from = key,
              values_from = value) |> 
  select(gdp, saving = invest) |> 
  mutate(year = lubridate::year(index)) |> 
  index_by(year) |> 
  group_by(year) |> 
  summarize(gdp = mean(gdp), saving = mean(saving)) |> 
  update_tsibble(key = c(gdp, saving), index = year)
```

There was a recession in the US economy in the year 1982. Therefore, we shall assume one structural break in the year 1982.

```{r}
yearly_us_macro |> 
  ggplot(aes(year, saving))+
  geom_line()+
  geom_line(aes(y = gdp))
```
