---
title: "Regression Models"
author: "Pravesh"
date: "September 2021"
output: 
  ioslides_presentation:
    logo: logo.png
    transition: slower
editor_options: 
  chunk_output_type: console
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 150px;
}
```

```{r, echo=FALSE}
options(scipen = 999, digits = 4)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```

```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(gapminder)
```



## Data

```{r, message=FALSE, warning=FALSE}
library(AER)
library(tidyverse)
data(Journals, package = "AER")
journals <- Journals %>% 
  mutate(citeprice = price/citations) %>% 
  select(subs, price, citeprice)
glimpse(journals)
```

## Summary

```{r}
library(summarytools)
dfSummary(journals, graph.col = FALSE ) 
```

## Regression model

- Estimate the effect of price per citation $citeprice$ on the number of subscriptions $subs$.
- Fit the following linear regression model
$$
\log(subs)_i = \beta_1 + \beta_2\log(citeprice)_i+ \epsilon_i
$$

## Plot {.smaller}
```{r}
library(ggplot2)
ggplot(data = journals, aes(log(subs), log(citeprice)))+
  geom_point() +
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) 
```


## Regression summary

```{r}
jour_lm <- lm(log(subs) ~ log(citeprice), data = journals)
summary(jour_lm)
```

## `lm` object
```{r}
class(jour_lm)
names(jour_lm)
jour_lm$coefficients
```


## Analysis of variance
```{r}
anova(jour_lm)
```

## Point and interval estimates

```{r}
# extract the estimated regression coefficients
coef(jour_lm)
```

```{r}
# calculate the confidence intervals
confint(jour_lm, level = 0.95)
```

## Testing a linear hypothesis {.smaller}

- Example: Test the hypothesis that the elasticity of number of subscriptions with respect to the price per citation equals $-0.5$, i.e. $H_0:\beta_2=-0.5$.

```{r}
library(car) # loads with AER
linearHypothesis(jour_lm, "log(citeprice) = -0.5")
```
- Cannot reject $H_0$.



## Data

```{r}
data(CPS1988, package = "AER")
# ?CPS1988
# summary(CPS1988)
dfSummary(CPS1988, graph.col = FALSE)

```


## Model
- The regression model we are interested is
$$
\log(wage) = \beta_1+ \beta_2experience+ \beta_3experience^2+ \\ \beta_4education+ \beta_5ethnicity + \epsilon
$$




---

```{r}
cps_lm <- lm(log(wage) ~ experience + I(experience^2) + education + ethnicity, data = CPS1988)
summary(cps_lm)
```


## Changing the reference category

```{r}
CPS1988$ethnicity <- relevel(CPS1988$ethnicity, ref = "afam")
summary(lm(log(wage) ~ experience + I(experience^2) + education + ethnicity, data = CPS1988))
```


## Comparison of models {.smaller}
- Is there any difference in the $\log(wage)$ (controlling for $experience$ and $education$) between Caucasian and African-American men?
- We already have a model `cps_lm` with all the variables including $ethnicity$.
- Now we need a model without $ethnicity$.
```{r}
noeth_lm <- update(cps_lm, formula =  . ~ . - ethnicity)
summary(noeth_lm)
```


## Comparison of models

```{r}
anova(cps_lm, noeth_lm )
```
The effect of $ethnicity$ is significant at any reasonable level.


## Interactions

```{r, echo=FALSE}
CPS1988$ethnicity <- relevel(CPS1988$ethnicity, ref = "cauc")
```


```{r}
cps_int <- lm(log(wage) ~ experience + I(experience^2) +
education * ethnicity, data = CPS1988)
# or cps_int <- lm(log(wage) ~ experience + I(experience^2) + education + ethnicity + education:ethnicity, data = CPS1988)
summary(cps_int)
```


## Separate regressions for each levels

```{r}
summary(
  update(cps_lm, formula = . ~ . - ethnicity, subset = ethnicity == "afam")
)
```


---


```{r}
summary(
  update(cps_lm, formula = . ~ . - ethnicity, subset = ethnicity == "cauc")
)
```